---
title: "Adult Classification Project"
author: Kevin Torres
output: html_document
---

**Task:** I will attempt to predict if peple in the data set belong in a certain class based on salary. (either making <= 50k or > 50k)

**Dataset:** I will be using a dataset with information about adults including 15 variables. 

Load csv file:

```{r setup, include=TRUE}
df = read.csv("adult_sal.csv")
adult <- as.data.frame(df)
head(adult)
library(dplyr)
```

Remove the first column and show head, srt, and summary to get an idea of the data

```{r}
adult <- select(adult, -X)
head(adult)
str(adult)
summary(adult)
```

Clean the data - Find categories to combine to simplify the data

```{r}
table(adult$type_employer)

# Combine the employer type 

unemp <- function(job){
  job <- as.character(job)
  if (job == 'Never-worked' | job == 'Without-pay') {
    return('Unemployed')
  } else{
    return(job)
  }
}

adult$type_employer <- sapply(adult$type_employer, unemp)
table(adult$type_employer)
```

Do the same for Job

```{r}
# Local and State level jobs

sl_gov <- function(job){
  job <- as.character(job)
  if (job == 'Local-gov' | job == 'State-gov') {
    return('SL-gov')
  } else{
    return(job)
  }
}

adult$type_employer <- sapply(adult$type_employer, sl_gov)

table(adult$type_employer)

# Self-employed

self_emp <- function(job){
  job <- as.character(job)
  if (job == 'Self-emp-inc' | job == 'Self-emp-not-inc') {
    return('Self-emp')
  } else{
    return(job)
  }
}

adult$type_employer <- sapply(adult$type_employer, self_emp)

table(adult$type_employer)
```

Do the same for Martial Status

```{r}
not_married <- function(status){
  status <- as.character(status)
  if (status == 'Divorced' | status == 'Separated' | status == 'Widowed') {
    return('Not-Married')
  } else{
    return(status)
  }
}

adult$marital <- sapply(adult$marital, not_married)

table(adult$marital)

married <- function(status){
  status <- as.character(status)
  if (status == 'Married-spouse-absent' | status == 'Married-AF-spouse' | status == 'Married-civ-spouse') {
    return('Married')
  } else{
    return(status)
  }
}

adult$marital <- sapply(adult$marital, married)

table(adult$marital)
```

Clean data - Country column

```{r}
# Group the countries by continent
Asia <- c('China', 'Hong', 'India', 'Iran', 'Cambodia', 'Japan', 'Laos', 'Philipines', 'Vietnam', 'Taiwan', 'Thailand')
North.America <- c('Canada', 'United-States', 'Puerto-Rico')
Europe <- c('England', 'France', 'Germany', 'Greece', 'Holand-Netherlands', 'Hungary', 'Ireland', 'Italy', 'Poland', 'Portugal', 'Scotland', 'Yugoslavia')
Latin.and.South.America <- c('Columbia', 'Cuba', 'Dominican-Republic', 'Ecuador', 'El-Salvador', 'Guatemala', 'Haiti', 'Honduras', 'Mexico', 'Nicaragua', 'Outlying-US(Guam-USVI-etc)','Peru', 'Jamaica', 'Trinidad&Tobago')
Other <- c('South')

group_country <- function(ctry){
  if (ctry %in% Asia) {
    return('Asia')
  } else if(ctry %in% North.America){
    return('North.America')
  } else if(ctry %in% Europe){
    return('Europe')
  } else if(ctry %in% Latin.and.South.America){
    return('Latin.and.South.America')
  } else{
    return('Other')
  }
}

adult$country <- sapply(adult$country, group_country)

table(adult$country)
```

Rename country column 
```{r}
# Now make each of the cleaned columns factors

adult$type_employer <- factor(adult$type_employer)
adult$marital <- factor(adult$marital)
adult$country <- factor(adult$country)

adult <- rename(adult, region = country)
str(adult)
```

Now that we have cleaned some of the data, lets check for NA values 

```{r}
library(Amelia)

# First lets convert the responses with '?' to an NA response
adult[adult == '?'] <- NA
str(adult)
table(adult$type_employer)
```

Now we can see the missing data on the map

```{r}
missmap(adult, main = 'Missingness Map', col = c("Yellow", 'Black'), legend = F)

# Remove NA values, even though this is not necessarily the best thing to do in most cases

adult <- na.omit(adult)

# Now that the data is clean, further explore it using visualization

str(adult)
missmap(adult, main = 'Missingness Map', col = c("Yellow", 'Black'), legend = F)
```

Exploratory Data Analysis 

```{r}
library(ggplot2)
# Histogram of ages, colored by income

ggplot(adult, aes(x = age, fill = income)) + geom_histogram(binwidth = 1, color = 'black') + theme_bw()

# Histogram of hours worked per week

ggplot(adult, aes(x = hr_per_week)) + geom_histogram(color = 'black', fill = 'light grey') + theme_bw()

# Barplot of region

ggplot(adult, aes(y = region, fill = income)) + geom_bar(color = 'black')
```

Now it's time to create a train and test data set

```{r}
head(adult)
library(caTools)

sample <- sample.split(adult$income, SplitRatio = .7)
train <- subset(adult, sample == T)
test <- subset(adult, sample == F)
```

Build the Logistic Regression Model

```{r}
model <- glm(income ~ ., family = binomial(link = 'logit'), data = train)
summary(glm(income ~ ., family = binomial(link = 'logit'), data = train))
# Here we see a lot of significant factors so there might be more factor groups we could create
```

Try to remove predictive variables from the model to attempt to delete variables that do not significantly add to the fit (use step function)

```{r}
new.step.model <- step(model)
summary(new.step.model)
```

Create a confuision matrix based on predicted values

```{r}
test$predicted.income <- predict(model, newdata = test, type = 'response')
table(test$income, test$predicted.income > 0.5)
```

Get measurements based off confusion matrix

```{r}
# Accuracy 
acc <- (6396+1409)/(6396+524+886+1409)
acc
# Recall
rcl <- 6396/(6396+524)
rcl
# Precision
prc <- 6396/(6396+886)
prc
```

We can determine if this is a good model after we find out the cost associated with accuracy, precision, and recall. Some models want to maximize one over the other. 








