---
title: "Bank Authentication Project"
author: "Kevin Torres"
output: 
  html_document: 
    keep_md: true
  
---

**Task:** I will be using Wavelet transformation data to see if a bank note is authentic or fake by building a neural net and making predictions based off it. I will then compare accuracy using a random forest model.

**Dataset:** The dataset consists of statistical information of images, therefore exploratory data analysis of this data is not easily interpretable.

Load csv file:

```{r setup, include=TRUE}
bank <- read.csv('bank_note_data.csv')
head(bank)
str(bank) 
```

Load necessary libraries and create train and test data sets

```{r}
library(caTools)

split <- sample.split(bank$Class, SplitRatio = .7)
train <- subset(bank, split == T)
test <- subset(bank, split == F)
```

Build the Neural Net 

```{r echo=TRUE}
library(neuralnet)
nn.bank <- neuralnet(Class ~ Image.Var + Image.Skew + Image.Curt + Entropy, data = train, hidden = 10, linear.output = F )
plot(nn.bank, rep = 'best')
```

Predictions

```{r}
predicted.nn.values <- compute(nn.bank, test[,1:4])
str(predicted.nn.values)
head(predicted.nn.values$net.result)
```

Here we notice the net results are still probabilities and we could use the round function to fix this

```{r}
predictions <- sapply(predicted.nn.values$net.result, round)
head(predictions)
```

Create a confusion matrix to see how we predicted 

```{r}
table(predictions, test$Class)

# We should be suspicious of perfect results since we did not even normalize the data
# We would typically normalize our data if there is a large range of min and max values between the column features
```

##############

Compare to a Random Forest Model

Load necessary libraries and create train and test data sets

```{r}
library(randomForest)

# First we need to set the Class column to be a factor, not an int like neural nets
bank$Class <- factor(bank$Class)

split <- sample.split(bank$Class, SplitRatio = .7)
train <- subset(bank, split == T)
test <- subset(bank, split == F)
```

Build the model

```{r}
nn.rf.model <- randomForest(Class ~ ., train)
nn.rf.model$confusion
```

Now we can predict
```{r}
rf.model.predict <- predict(nn.rf.model, test)
table(rf.model.predict, test$Class)

# This model was almost perfect, therefore we can conclude that we should not be suspicious of our perfect neural net model 
```
