---
title: "Lending Club Project"
author: Kevin Torres
output: html_document
---

**Task:** I will create a model to help predict people who showed a profile of having a high probability of paying investors back. I will try to classify and predict whether or not the borrower paid their loan back in full.

**Dataset:** I will be using publicly available lending data from 2007-2010. This data is from Lending Club which connects people who need money with investors. 

Load the .csv file and save it as a data frame called loans

```{r setup, include=TRUE}
loans <- read.csv('loan_data.csv')
loans <- as.data.frame(loans)
head(loans)
str(loans)
summary(loans)
```

Convert certain columns to categorical data (factor)

```{r}
loans$inq.last.6mths <- factor(loans$inq.last.6mths)
loans$delinq.2yrs <- factor(loans$delinq.2yrs)
loans$pub.rec <- factor(loans$pub.rec)
loans$not.fully.paid <- factor(loans$not.fully.paid)
loans$credit.policy <- factor(loans$credit.policy)

str(loans)
```

Exploratory Data Analysis

```{r}
library(ggplot2)

# Histogram of fico scores colored by whether the loan was fully paid or not
ggplot(loans, aes(fico, fill = not.fully.paid)) + geom_histogram(color = 'black') + theme_bw()

# Barplot of the purpose of loan counts, colored by whether the loan was fully paid or not
ggplot(loans, aes(x = factor(purpose), fill = not.fully.paid)) + geom_bar(color ='black', position = position_dodge()) + theme_bw()

# Scatterplot showing fico score versus interest rate 
ggplot(loans, aes(int.rate,fico, color = not.fully.paid)) + geom_point(alpha = .4) + theme_bw()
```

Split data into train and test sets

```{r}
library(caTools)

sample <- sample.split(loans$not.fully.paid, SplitRatio = .7)
train <- subset(loans, sample == T)
test <- subset(loans, sample == F)
```

Build and train a model using the svm function

```{r}
library(e1071)

model <- svm(not.fully.paid ~ ., train)
summary(model)
```

Only pass in test data that does not have a label 

```{r}
predicted.values <- predict(model, test[1:13])
table(predicted.values, test$not.fully.paid)

# These are bad results, model is predicting that everyone paid off their loan
# We need to use the tune function to try out different cost and gamma values
```

Use the tune function to try out different cost and gamma values


```{r}
tuned.results <- tune(svm, train.x = not.fully.paid ~ ., data = train, kernel = 'radial', ranges = list(cost = c(100,125,150,200), gamma= c(.075,.1)))
summary(tuned.results)

# We find our best parameters are 100 and .075
tuned.model <- svm(not.fully.paid ~ ., data = train, cost = 100, gamma = .075)
```

Make predictions based off this tuned model

```{r}
tuned.predictions <- predict(tuned.model, test[1:13])
table(tuned.predictions, test$not.fully.paid)

# Now we are no longer classifying everyone as paying off their loan
```

This model still is not too great but we could add more values to the ranges for cost and gamma but it would take time 


